---
title: "Diagnostic Forestier Nemeton"
subtitle: "Rapport de synthese"
lang: en
date: today
format:
  pdf:
    documentclass: article
    papersize: a4
    fontsize: 11pt
    geometry:
      - top=25mm
      - bottom=25mm
      - left=20mm
      - right=20mm
    colorlinks: true
    linkcolor: "ForestGreen"
    toc: false
    number-sections: true
    fig-pos: "H"
    include-in-header:
      text: |
        \usepackage{eso-pic}
        \usepackage{tikz}
        \usepackage{tcolorbox}
        \usepackage{colortbl}
        \usepackage{xcolor}
        \usepackage{booktabs}
        \usepackage{longtable}
        \tcbuselibrary{skins}
params:
  data_file: ""
  radar_file: ""
  family_maps_dir: ""
  language: "fr"
  cover_image: ""
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false

# Load report data
report_data <- readRDS(params$data_file)

# Helper for translations
tr <- function(key) {
  report_data$labels[[key]] %||% key
}

# Color formatting helper (returns LaTeX color name)
score_color <- function(score) {
  if (is.na(score)) return("gray")
  if (score >= 60) "ForestGreen"
  else if (score >= 40) "Orange"
  else "Red"
}

# Quality label helper
quality_label <- function(score, lang = params$language) {
  if (is.na(score)) return("N/A")
  if (lang == "fr") {
    if (score >= 80) "Excellent"
    else if (score >= 60) "Bon"
    else if (score >= 40) "Moyen"
    else if (score >= 20) "Faible"
    else "Tr\u00e8s faible"
  } else {
    if (score >= 80) "Excellent"
    else if (score >= 60) "Good"
    else if (score >= 40) "Average"
    else if (score >= 20) "Low"
    else "Very low"
  }
}

# Quality color for LaTeX (hex)
quality_color_hex <- function(score) {
  if (is.na(score)) return("808080")
  if (score >= 80) "228B22"
  else if (score >= 60) "32CD32"
  else if (score >= 40) "FF8C00"
  else if (score >= 20) "DC143C"
  else "8B0000"
}

# Convert hex color to LaTeX \definecolor-compatible format
hex_to_latex_rgb <- function(hex) {
  hex <- sub("^#", "", hex)
  r <- strtoi(substr(hex, 1, 2), 16L) / 255
  g <- strtoi(substr(hex, 3, 4), 16L) / 255
  b <- strtoi(substr(hex, 5, 6), 16L) / 255
  sprintf("%.3f,%.3f,%.3f", r, g, b)
}

# Escape LaTeX special characters
escape_latex <- function(x) {
  x <- gsub("\\\\", "\\\\textbackslash{}", x)
  x <- gsub("([&%$#_{}])", "\\\\\\1", x)
  x <- gsub("~", "\\\\textasciitilde{}", x)
  x <- gsub("\\^", "\\\\textasciicircum{}", x)
  x
}

# Family order
family_order <- c("C", "B", "W", "A", "F", "L", "T", "R", "S", "P", "E", "N")

# Sanitize markdown text to prevent YAML parse issues
sanitize_md <- function(text) {
  if (is.null(text) || nchar(text) == 0) return(text)
  text <- gsub("\n---\n", "\n***\n", text)
  text <- gsub("^---$", "***", text)
  text
}
```

```{r cover-page}
#| results: asis

# ---- Cover page ----
cat("\\thispagestyle{empty}\n")
cat("\\newgeometry{top=0mm,bottom=0mm,left=0mm,right=0mm}\n")

has_cover <- nchar(params$cover_image) > 0 && file.exists(params$cover_image)

if (has_cover) {
  # Normalize path for LaTeX (forward slashes, no backslashes)
  cover_path <- normalizePath(params$cover_image, winslash = "/", mustWork = FALSE)
  # Cover with background image
  cat("\\begin{tikzpicture}[remember picture, overlay]\n")
  cat("  \\node[anchor=center, inner sep=0pt] at (current page.center) {\n")
  cat("    \\includegraphics[width=\\paperwidth, height=\\paperheight, keepaspectratio=false]{",
      cover_path, "}\n", sep = "")
  cat("  };\n")
  # Semi-transparent white overlay at top
  cat("  \\fill[white, opacity=0.88] (current page.north west) rectangle ",
      "([yshift=-7cm]current page.north east);\n")
  # Semi-transparent white overlay at bottom
  cat("  \\fill[white, opacity=0.88] ([yshift=6cm]current page.south west) rectangle ",
      "(current page.south east);\n")
  cat("\\end{tikzpicture}\n\n")
}

# Title block at top
cat("\\vspace*{1cm}\n")
cat("\\begin{center}\n")
cat("{\\fontsize{28}{34}\\selectfont\\textcolor[HTML]{2E7D32}{\\textbf{",
    escape_latex(tr("title")), "}}}\n\n", sep = "")
cat("\\vspace{0.3cm}\n")
cat("{\\Large\\textcolor{gray}{",
    escape_latex(tr("subtitle")), "}}\n\n", sep = "")
cat("\\vspace{1cm}\n")
cat("{\\LARGE\\textbf{", escape_latex(report_data$project_name), "}}\n\n", sep = "")
cat("\\end{center}\n\n")

# Score block at bottom
cat("\\vfill\n")
cat("\\begin{center}\n")
score_col <- quality_color_hex(report_data$global_score)
cat("\\textcolor[HTML]{", score_col, "}{\\fontsize{48}{58}\\selectfont\\textbf{",
    sprintf("%.1f", report_data$global_score), "}}\n\n", sep = "")
cat("{\\Large / 100}\n\n")
cat("\\vspace{0.3cm}\n")
# NDP badge sur la page de couverture
ndp_col <- switch(as.character(report_data$ndp_level),
  "0" = "6c757d", "1" = "17a2b8", "2" = "28a745",
  "3" = "ffc107", "4" = "fd7e14", "6c757d")
confidence_pct <- round(report_data$ndp_confidence * 100, 1)
cat("\\colorbox[HTML]{", ndp_col, "}{\\textcolor{white}{\\small\\textbf{ NDP ",
    report_data$ndp_level, " -- ", escape_latex(report_data$ndp_name), " }}}\n\n", sep = "")
cat("{\\small ", escape_latex(tr("ndp_confidence_label")),
    " : ", confidence_pct, "\\%}\n\n", sep = "")
cat("\\vspace{0.3cm}\n")
cat("{\\large ", report_data$n_ugs, " UGF / ",
    report_data$n_parcels, " ",
    escape_latex(tr("parcels")), "}\n\n", sep = "")
cat("\\vspace{0.8cm}\n")
cat("{\\small\\textcolor{gray}{", escape_latex(tr("generated_by")),
    " v", as.character(packageVersion("nemeton")),
    " --- ", as.character(Sys.Date()), "}}\n", sep = "")
cat("\\end{center}\n\n")
cat("\\restoregeometry\n")
cat("\\newpage\n")
```

# `r tr("project_info")`

**`r tr("project_name")`:** `r report_data$project_name`

`r if (!is.null(report_data$project_description) && nchar(report_data$project_description) > 0) paste0("**Description:** ", report_data$project_description)`

`r if (!is.null(report_data$project_owner) && nchar(report_data$project_owner) > 0) paste0("**", tr("owner"), ":** ", report_data$project_owner)`

- **UGF :** `r report_data$n_ugs`
- **`r tr("parcels")`:** `r report_data$n_parcels`
- **`r tr("created_at")`:** `r report_data$created_at`

\begin{center}
\textcolor{`r score_color(report_data$global_score)`}{\LARGE\textbf{`r tr("global_score_label")` : `r sprintf("%.1f", report_data$global_score)` / 100}}

\vspace{0.2cm}
{\small `r tr("ndp_label")` : \textbf{NDP `r report_data$ndp_level` -- `r report_data$ndp_name`} \quad | \quad `r tr("ndp_confidence_label")` : `r sprintf("%.1f", report_data$ndp_confidence * 100)`\%}
\end{center}

\vspace{0.3cm}

```{r radar}
#| out-width: "95%"
#| fig-align: center

if (file.exists(params$radar_file)) {
  knitr::include_graphics(params$radar_file)
} else {
  cat("Radar plot not available")
}
```

\newpage

# `r tr("family_scores")`

```{r family-table}
#| results: asis

# Build colorized LaTeX table
cat("\\begin{center}\n")
cat("\\begin{tabular}{clccc}\n")
cat("\\toprule\n")

# Header
if (params$language == "fr") {
  cat("\\textbf{Code} & \\textbf{Famille} & \\textbf{Score} & \\textbf{Min} & \\textbf{Max} \\\\\n")
} else {
  cat("\\textbf{Code} & \\textbf{Family} & \\textbf{Score} & \\textbf{Min} & \\textbf{Max} \\\\\n")
}
cat("\\midrule\n")

for (code in family_order) {
  if (!code %in% names(report_data$family_stats)) next
  fam <- report_data$family_stats[[code]]
  fam_hex <- sub("^#", "", fam$color)

  cat("\\textcolor[HTML]{", fam_hex, "}{\\textbf{", escape_latex(fam$code), "}} & ",
      escape_latex(fam$name), " & ",
      "\\textcolor[HTML]{", fam_hex, "}{\\textbf{", sprintf("%.1f", fam$mean), "}} & ",
      sprintf("%.1f", fam$min), " & ",
      sprintf("%.1f", fam$max), " \\\\\n", sep = "")
}

cat("\\bottomrule\n")
cat("\\end{tabular}\n")
cat("\\end{center}\n")
```

\vspace{1cm}

## `r if (params$language == "fr") "Interpr\u00e9tation des scores" else "Score interpretation"`

| Score | `r if (params$language == "fr") "Interpr\u00e9tation" else "Interpretation"` |
|-------|----------------|
| 80-100 | `r if (params$language == "fr") "Excellent - Service \u00e9cosyst\u00e9mique tr\u00e8s performant" else "Excellent - Very high ecosystem service performance"` |
| 60-79 | `r if (params$language == "fr") "Bon - Service \u00e9cosyst\u00e9mique performant" else "Good - High ecosystem service performance"` |
| 40-59 | `r if (params$language == "fr") "Moyen - Service \u00e9cosyst\u00e9mique moyen" else "Average - Moderate ecosystem service performance"` |
| 20-39 | `r if (params$language == "fr") "Faible - Service \u00e9cosyst\u00e9mique limit\u00e9" else "Low - Limited ecosystem service performance"` |
| 0-19 | `r if (params$language == "fr") "Tr\u00e8s faible - Service \u00e9cosyst\u00e9mique d\u00e9ficient" else "Very low - Deficient ecosystem service performance"` |

```{r}
#| label: synthesis-comments
#| results: asis

# Render synthesis comments if available
if (!is.null(report_data$synthesis_comments) && nchar(report_data$synthesis_comments) > 0) {
  cat("\n\\newpage\n\n")
  cat("# ", tr("comments_title"), "\n\n", sep = "")
  cat(sanitize_md(report_data$synthesis_comments))
  cat("\n\n")
}
```

\newpage

```{r toc}
#| results: asis
cat("\\tableofcontents\n")
```

\newpage

# `r if (params$language == "fr") "D\u00e9tail par famille d'indicateurs" else "Detail by indicator family"`

```{r family-sections}
#| results: asis

for (code in family_order) {
  if (!code %in% names(report_data$family_stats)) next

  fam <- report_data$family_stats[[code]]
  fam_hex <- sub("^#", "", fam$color)

  # Colored section header using tcolorbox
  cat("\n\n")
  cat("\\begin{tcolorbox}[colback=white, colframe={rgb,255:red,",
      strtoi(substr(fam_hex, 1, 2), 16L), ";green,",
      strtoi(substr(fam_hex, 3, 4), 16L), ";blue,",
      strtoi(substr(fam_hex, 5, 6), 16L), "},",
      " colbacktitle={rgb,255:red,",
      strtoi(substr(fam_hex, 1, 2), 16L), ";green,",
      strtoi(substr(fam_hex, 3, 4), 16L), ";blue,",
      strtoi(substr(fam_hex, 5, 6), 16L), "},",
      " coltitle=white, title={\\Large\\textbf{",
      escape_latex(fam$name), " (", code, ")}},",
      " fonttitle=\\sffamily, boxrule=0.5mm, arc=2mm]\n", sep = "")

  # Score summary inside the box
  score_label <- if (params$language == "fr") "Score moyen" else "Mean score"
  cat("\\textbf{", escape_latex(score_label), " :} ",
      "\\textcolor[HTML]{", fam_hex, "}{\\textbf{\\Large ",
      sprintf("%.1f", fam$mean), " / 100}}", sep = "")
  cat(" \\hfill (Min: ", sprintf("%.1f", fam$min),
      ", Max: ", sprintf("%.1f", fam$max), ")\n\n", sep = "")

  # Family description inside the box
  if (!is.null(report_data$family_descriptions[[code]])) {
    cat("\\smallskip\n")
    cat(escape_latex(report_data$family_descriptions[[code]]), "\n")
  }

  cat("\\end{tcolorbox}\n\n")

  # Map image if available
  map_file <- file.path(params$family_maps_dir, paste0("family_", code, "_map.png"))
  if (file.exists(map_file)) {
    cat("### ", if (params$language == "fr") "Carte" else "Map", "\n\n", sep = "")
    rel_path <- file.path("family_maps", paste0("family_", code, "_map.png"))
    cat("![](", rel_path, "){width=90%}\n\n", sep = "")
  }

  # Indicator details table with colored scores
  if (!is.null(report_data$indicator_stats[[code]])) {
    ind_data <- report_data$indicator_stats[[code]]
    if (nrow(ind_data) > 0) {
      cat("### ", if (params$language == "fr") "Indicateurs" else "Indicators", "\n\n", sep = "")

      cat("\\begin{center}\n")
      cat("\\begin{tabular}{lccc}\n")
      cat("\\toprule\n")
      if (params$language == "fr") {
        cat("\\textbf{Indicateur} & \\textbf{Moyenne} & \\textbf{Min} & \\textbf{Max} \\\\\n")
      } else {
        cat("\\textbf{Indicator} & \\textbf{Mean} & \\textbf{Min} & \\textbf{Max} \\\\\n")
      }
      cat("\\midrule\n")

      for (i in seq_len(nrow(ind_data))) {
        score_hex <- quality_color_hex(ind_data[i, 2])
        cat(escape_latex(as.character(ind_data[i, 1])), " & ",
            "\\textcolor[HTML]{", score_hex, "}{\\textbf{", sprintf("%.1f", ind_data[i, 2]), "}} & ",
            sprintf("%.1f", ind_data[i, 3]), " & ",
            sprintf("%.1f", ind_data[i, 4]), " \\\\\n", sep = "")
      }

      cat("\\bottomrule\n")
      cat("\\end{tabular}\n")
      cat("\\end{center}\n\n")
    }
  }

  # Per-UGF scores table
  if (!is.null(report_data$parcel_scores[[code]])) {
    ps <- report_data$parcel_scores[[code]]
    if (nrow(ps) > 0) {
      cat("### ", if (params$language == "fr") "Scores par UGF" else "Scores per UGF", "\n\n", sep = "")

      cat("\\begin{center}\n")
      cat("\\begin{tabular}{lcc}\n")
      cat("\\toprule\n")
      if (params$language == "fr") {
        cat("\\textbf{UGF} & \\textbf{Score} & \\textbf{Qualit\u00e9} \\\\\n")
      } else {
        cat("\\textbf{UGF} & \\textbf{Score} & \\textbf{Quality} \\\\\n")
      }
      cat("\\midrule\n")

      n_show <- min(nrow(ps), 20L)
      for (i in seq_len(n_show)) {
        val <- ps$score[i]
        q_hex <- quality_color_hex(val)
        q_label <- quality_label(val)
        score_str <- if (is.na(val)) "N/A" else sprintf("%.1f", val)

        cat(escape_latex(ps$id[i]), " & ", score_str, " & ",
            "\\textcolor[HTML]{", q_hex, "}{\\textbf{", escape_latex(q_label), "}} \\\\\n", sep = "")
      }

      cat("\\bottomrule\n")
      cat("\\end{tabular}\n")
      cat("\\end{center}\n\n")

      # Note if truncated
      if (nrow(ps) > 20) {
        remaining <- nrow(ps) - 20
        msg <- if (params$language == "fr") {
          sprintf("\\textit{... %d UGF suppl\u00e9mentaires}", remaining)
        } else {
          sprintf("\\textit{... %d additional UGFs}", remaining)
        }
        cat("\\begin{center}\n", msg, "\n\\end{center}\n\n", sep = "")
      }
    }
  }

  # Family comments (markdown, sanitized)
  if (!is.null(report_data$family_comments[[code]]) &&
      nchar(report_data$family_comments[[code]]) > 0) {
    cat("### ", if (params$language == "fr") "Analyse" else "Analysis", "\n\n", sep = "")
    cat(sanitize_md(report_data$family_comments[[code]]), "\n\n")
  }

  cat("\n\\newpage\n")
}
```

```{r ug-composition, results='asis'}
#| label: ug-composition
# UG Composition section (only if UG data is available)
if (isTRUE(report_data$has_ugs) && !is.null(report_data$ug_composition) &&
    length(report_data$ug_composition) > 0) {

  cat("# ", tr("ug_section_title"), "\n\n", sep = "")

  if (params$language == "fr") {
    cat(sprintf("Le projet comporte **%d unit\u00e9(s) de gestion** r\u00e9parties sur %d parcelle(s) cadastrale(s).\n\n",
                report_data$n_ugs, report_data$n_parcels))
  } else {
    cat(sprintf("The project contains **%d management unit(s)** across %d cadastral parcel(s).\n\n",
                report_data$n_ugs, report_data$n_parcels))
  }

  # Summary table
  cat("\\begin{center}\n")
  cat("\\begin{tabular}{llrl}\n")
  cat("\\toprule\n")
  if (params$language == "fr") {
    cat("\\textbf{UG} & \\textbf{Groupe} & \\textbf{Surface (ha)} & \\textbf{R\u00e9f\u00e9rences cadastrales} \\\\\n")
  } else {
    cat("\\textbf{UG} & \\textbf{Group} & \\textbf{Area (ha)} & \\textbf{Cadastral references} \\\\\n")
  }
  cat("\\midrule\n")

  for (ug in report_data$ug_composition) {
    groupe_str <- if (is.null(ug$groupe) || is.na(ug$groupe)) "---" else ug$groupe
    cat(
      escape_latex(ug$label), " & ",
      escape_latex(groupe_str), " & ",
      sprintf("%.2f", ug$surface_ha), " & ",
      escape_latex(ug$cadastral_refs), " \\\\\n",
      sep = ""
    )
  }

  cat("\\bottomrule\n")
  cat("\\end{tabular}\n")
  cat("\\end{center}\n\n")

  # Detailed annex: ONF Annexe 2 style (per-UG cadastral breakdown)
  if (params$language == "fr") {
    cat("## Annexe : Composition cadastrale d\u00e9taill\u00e9e\n\n")
  } else {
    cat("## Annex: Detailed Cadastral Composition\n\n")
  }

  for (ug in report_data$ug_composition) {
    groupe_str <- if (is.null(ug$groupe) || is.na(ug$groupe)) "" else paste0(" [", ug$groupe, "]")
    cat(sprintf("**%s**%s --- %.2f ha\n\n",
                escape_latex(ug$label), escape_latex(groupe_str), ug$surface_ha))

    refs <- ug$refs_detail
    if (!is.null(refs) && is.data.frame(refs) && nrow(refs) > 0) {
      cat("\\begin{center}\n")
      cat("\\begin{tabular}{llrr}\n")
      if (params$language == "fr") {
        cat("\\textbf{R\u00e9f. cadastrale} & \\textbf{Section} & \\textbf{N\\textsuperscript{o}} & \\textbf{Surface (m\\textsuperscript{2})} \\\\\n")
      } else {
        cat("\\textbf{Cadastral ref.} & \\textbf{Section} & \\textbf{No.} & \\textbf{Area (m\\textsuperscript{2})} \\\\\n")
      }
      cat("\\hline\n")

      for (j in seq_len(nrow(refs))) {
        cat(
          escape_latex(as.character(refs$geo_parcelle[j])), " & ",
          escape_latex(as.character(refs$section[j])), " & ",
          escape_latex(as.character(refs$numero[j])), " & ",
          format(round(refs$surface_m2[j]), big.mark = " "), " \\\\\n",
          sep = ""
        )
      }

      cat("\\end{tabular}\n")
      cat("\\end{center}\n\n")
    }
  }

  cat("\\newpage\n")
}
```

\vfill

*`r tr("generated_by")` v`r packageVersion("nemeton")` - `r Sys.Date()`*
